Structuring College Access: The Market Segment Model and College Board Geomarkets

### 2025-05-14 text/back_end_KS.qmd
dallas revisions

<!--# DISCUSSION OF THIS FIGURE SOMEWHAT BURIES THE LEAD THAT CHICAGO LOOKS TOTALLY DIFFERENT THAN THE OTHERS IN TERMS OF RACIAL COMPOSITION OF PURCHASED PROFILES! --> @fig-chicago-rq2-race presents the racial/ethnic composition of purchased student profiles within Geomarkets in the Chicago metropolitan area, notably revealing that the City of Chicago stands out distinctly from all other Geomarkets in terms of the substantial racial and ethnic diversity of its purchased profiles. At the lowest PSAT score thresholds (top panel, 1020-1150) the racial/ethnic distribution of prospects within Geomarkets most closely reflects the demographic makeup of the metropolitan area in 2020 (see @fig-chicago-rq1-race). For example, more than 8,000 of the 15,981 total prospects in this SAT range order were concentrated in the Western Suburbs (56% White, 8% Asian, 6% Black, 26% Hispanic, 4% Multiracial) and South and Southwest Suburbs (55% White, 3% Asian, 17% Black, 22% Hispanic, and 3% Multiracial), both of which yield majority White prospect pools comparable to the racial/ethnic makeup of their overall Geomarket populations. The North Shore (78% White, 8% Asian, 2% Black, 8% Hispanic) and the Northwest Suburbs (61% White, 11% Asian, 2% Black, 23% Hispanic) Geomarket pools exhibit the greatest proportions of White prospects. Unlike any other Geomarket, nearly nine of every 10 prospects from the City of Chicago Geomarket were Hispanic (50%), Black (30%), Asian (5%), or Multiracial (2%).

@@

[WRITING IS QUITE GOOD!I THINK THERE IS A MODEST LEAD THAT THE NEXT COUPLE PARAGRAPHS DON'T STRESS ENOUGH , WHICH IS THAT TX19/CITY OF DALLAS IS THE MOST LEAST WHITE/ASIAN AND THE POOREST]
@fig-dallas-rq1-race presents the racial and ethnic composition of each Geomarket within the Dallas-Fort Worth metropolitan area across the three census periods. In 1980, all Geomarkets were predominantly White, with White populations ranging from 62% in City of Dallas to nearly 92% in West of Dallas/Fort Worth Metroplex. By 2000, however, Geomarkets located in the central part of the metropolitan area exhibited a more balanced racial/ethnic distribution, including Dallas County (58% White, 16% Black, 19% Hispanic; 6% Asian), Irving, Arlington, and Grand Prairie (54% White, 12% Black, 26% Hispanic; 16% Asian), and City of Fort Worth (57% White, 16% Black, 22% Hispanic; 3% Asian). In contrast, Collin & Rockwall Counties and West of Dallas/Fort Worth Metroplex Geomarkets remained predominantly White in 2000 at 78% and 80%, respectively. Notably, by 2000, the City of Dallas was no longer a predominantly White Geomarket, with a racial/ethnic composition of approximately 24% Black, 30% Hispanic, and 3% Asian residents.

These shifts in racial and ethnic composition within and across Geomarkets became more pronounced between 2000 and 2020. The Collin and Rockwall and West of Dallas/Fort Worth Metroplex Geomarkets experienced continued declines in their share of White residents. However, both Geomarkets remained predominantly White (58% and 63%, respectively). In contrast, all other Geomarkets experienced continued increases in the shares of Black and Hispanic residents, resulting in a diminished previously White-majority populations for Dallas County (33% White, 23% Black, 32% Hispanic; 9% Asian), Irving, Arlington, and Grand Prairie (30% White, 20% Black, 36% Hispanic; 12% Asian), and City of Fort Worth (46% White, 15% Black, 33% Hispanic; 4% Asian). Lastly, continuous increases in the shares of Black and Hispanic residents further diversified the City of Dallas, making it the most racially/ethnically diverse Geomarket in the Dallas-Fort Worth metropolitan area.

[CITY OF DALLAS!]Socioeconomic demographics of median income, percent living below poverty, and educational attainment over time are presented in @fig-dallas-rq1-ses for the Dallas-Fort Worth metropolitan area. Overall, median income both within and across Geomarkets remained relatively stable across the three census periods. Between 1980 to 2000, affluence levels across Geomarkets were consistent, ranging from a low of \$82,000 in City of Fort Worth to a high of \$135,00 in the Collin & Rockwall Counties Geomarket. However, the substantial increase in median income for the Collin & Rockwall Counties Geomarket during this time period plateaued by 2020, reshuffling relatively affluence levels among Geomarkets. By 2020, Collin & Rockwall Counties remained the most affluent Geomarket with a median household income of \$133,000, followed by West of Dallas/Fort Worth Metroplex (\$116,000), and Dallas County (\$90,000). In contrast, median household income for City of Dallas and City of Fort Worth remained relatively stable and continued to rank at the lower levels of affluence within the metropolitan area. \[KS: NEED TO CHECK THIS; I THINK CITY OF DALLAS AND CITY OF FORT WORTH ARE MIXED UP BETWEEN GRAPH VERSUS TABLE\]

[DON'T KNOW IF WE NEED PARAGRAPH ON POVERTY + EDUCATIONAL ATTAINMENT]. @fig-dallas-rq1-ses also illustrates changes in the percentage of residents living in poverty within Geomarkets over time, revealing shifting dynamics across Geomarkets similar to those observed in median household income. In 1980, poverty rates were comparably low in both Collin & Rockwall Counties and Dallas County, moderate in the Irving, Arlington, and Grand Prarie and West of Dallas/Fort Worth Metroplex Geomarkets, and highest for City of Dallas and City of Fort Worth Geomarkets. Overtime, poverty rates declined in Collin & Rockwall Counties and West of Dallas/Fort Worth Metroplex, while increasing for all other Geomarkets. These divergent trends resulted in a widening gap in poverty levels across the Dallas-Fort Worth metropolitan area. A similar widening disparity is evident in educational attainment. Although all Geomarkets experienced increases in the percentage of residents with a bachelor's degree or higher between 1980 and 2020, the Collin & Rockwall Counties Geomarket experienced a substantially greater increase - from 23% in 1980 to 50% in 2020- maintaining the gap with other Geomarkets in terms of educational attainment.

[DON'T THINK YOU NEED SUMMARY PARAGRAPHS LIKE THIS WITHIN SECTIONS FOR THIS METRO]The Dallas-Fort Worth metroplex experienced substantial demographic growth since Geomarket boundaries were dilineated by @RN4982 in the early 1980s, leading to an urban landscape with considerable racial/ethnic diversity across multiple Geomarkets (as opposed to concentrated within a single Geomarket like Chicago) but stark socioeconomic disparities between Geomarkets. While all six Geomarkets experienced population increases, the most significant growth occurred in the outer suburban areas, reflecting ongoing urban sprawl. Such growth is largely attributed to increases in Hispanic, Black, and Asian populations leading to only two of the six Dallas-Fort Worth Geomarkets remaining predominantly White (Collin and Rockwall Counties and West of Dallas/Fort Worth Metroplex. On the other hand, the metropolitan area has experienced widening disparities in income, poverty, and education attainment between Geomarkets, especially between the historically affluent Collin and Rockwall Counties and the less economically advantaged inner-core Geomarkets (City of Fort Worth and City of Dallas). Together, these trends underscore the growingly racial/ethnic diverse but widening economic inequality within the evolving Geomarket landscape of the Dallas-Fort Worth metroplex.

@@

Contributions of each Geomarket to the pool of prospects based on first-generation college student status from the Dallas-Fort Worth metropolitan area are presented in @fig-dallas-rq2-midSAT-ses. Only 303 prospects of the 3,928 total in this middle PSAT score range pool were first-generation college students whose parents did not attend college. Again, all Geomarkets besides Collin and Rockwall Counties (14% versus 36% for all prospects) and West of Dallas/Fort Worth Metroplex (15% versus 24% for all prospects) contributed equal or larger proportions of first-generation college students (no college) relative to their contributions to the overall prospect pool. Disproportioantely larger contributions range from six percentage points by the Dallas Counties Geomarket (19% versus 13% for all prospects) to 11 percentage points by the City of Dallas Geomarket (22% versus 11% for all prospects). Contributions become more proportional for the pool of first-generation college students whose parents attended but did not complete their degree. Proportional balances across Geomarkets are also evident for the 3,069 prospects that are not first-generation college students, with the exception of Collin and Rockwall Counties Geomarket contributing the largest disproportinate share of not first-generation college students (40%) relative to the contributions to the overall prospect pool (36%).

@@

By contrast, for non-White groups, a larger share of prospects were first-generation students. Furthermore, first generation status differed substantially across Geomarkets, especially for Hispanic and Asian students. For Hispanic students, more than 50% of prospects from the City of Dallas, City of Forth Worth, Arlington \& Grand Prairie, and Dallas County are first-generation college students. By contrast, for the more affluent Collin & Rockwall Geomarket and the West of Dallas/Fort Worth Metroplex Geomarkets, less than 30% of prospects were first-generation. For Asian prospects, City of Fort Worth and Irving, Arlington, & Grand Prairie Geomarkets contributed disproportionately larger shares of first-generation college students across both parents with no college and some college. By contrast, the more affluent Collin & Rockwall sent more Asian prospects (285) than any other Geomarket, but about 87% of these students had parents with a BA.

<!-- I ADDED THE ABOVE PARAGRAPH AND CUT THIS PARAGRAPH TO GIVE THE LEDE MORE PUNCH 
Similar patterns of one or two Geomarkets disproportionately contributing larger shares of first-generation college students was evident across other racial/ethnic categories in @fig-dallas-rq2-midSAT-combo. For instance, City of Dallas (18%) and City of Fort Worth (24%) contributed larger proportions of first-generation college students whose parents did not attend college relative to the overall pool of all prospects (8%). On the other hand, the Irving, Arlington, & Grand Prairie (23%) and Dallas Counties (26%) Geomarkets contributed disproportionately larger proportions of Black students whose parents attended some college but did not complete their degree relative to the overall pool of all prospects (17%). For Asian prospects, City of Fort Worth and Irving, Arlington, & Grand Prairie Geomarkets contributed disproportionately larger shares of first-generation college students across both parents with no college and some college. City of Dallas (38%), City of Fort Worth (39%), Irving, Arlington, & Grand Prairie (32%), Dallas Counties (28%) all contributed larger shares of Hispanic students whose parents did not attend college relative to the overall prospect pool of these first-generation college students (23%), whereas only Dallas Counties contributed to larger shares of Hispanic students who parents attended some college but did not attain their degree (26% versus 20% for all prospects).
-->

Analyses of student list orders across the Dallas-Fort Worth metropolitan area reveals patterns of racial/ethnic and socioeconomic stratification in Geomarket contributions. As Dallas-Fort Worth became increasingly diverse from 1980 to 2020, a more balanced racial/ethnic distribution emerged across the metropolitan area's Geomarket landscape. However, these patterns did not translate to racial/ethnic composition of student prospect pools. Suburban Geomarkets like Collin and Rockwall Counties and the West of Dallas/Fort Worth Metroplex yielded a disproportionately larger share of Asian prospects, whereas central Geomarkets such as the City of Dallas, Dallas County, and Irving, Arlington, and Grand Prairie contributed a disproportionately larger share of Hispanic and Black prospects. These racial/ethnic disparities become significantly more pronounced, particularly in the disproportionate share of Asian prospects, as PSAT score thresholds increase (see Appendix @fig-dallas-rq2-lowSAT-race and @fig-dallas-rq2-highSAT-race). Similar Geomarket contribution patterns are also evident by first-generation college student status, with Collin and Rockwall Counties and the West of Dallas/Fort Worth Metroplex providing larger shares of non-first generation college student prospects, and central Geomarkets (City of Dallas, City of Fort Worth, Dallas County, and Irving, Arlington, and Grand Prairie) providing disproportionately larger shares of first-generation college student prospects. Moreover, central Geomarkets are particularly important for including Black (City of Dallas and City of Fort Worth), Hispanic (City of Dallas, City of Fort Worth, Dallas County, and Irving, Arlington, and Grand Prairie), and Asian (City of Fort Worth and Irving, Arlington, and Grand Prairie) first-generation college students. While Collin and Rockwall Counties and the West of Dallas/Fort Worth Metroplex Geomarkets still contribute a notable overall number of Black, Hispanic, and Asian prospects, such students are predominantly not first-generation college students.

@@

\@fig-la-rq1-race presents the racial and ethnic composition of each Geomarket within the Los Angeles metropolitan area from 1980 to 2020. Over this period, the proportion of White residents declined across all Geomarkets, while Asian and Hispanic populations generally increased. [THE CAVEAT THAT WE DON'T HAVE ASIAN FOR 1980 IS MORE IMPORTANT FOR LA] In 1980, all Geomarkets - except for East Los Angeles, South and South Central Los Angeles, and Hollywood and Wilshire- were predominantly White, with percentages ranging from 62% White in Long Beach to 78% in the San Fernando Valleys (West and East) Geomarkets. By 2020, however, only the West Los Angeles and West Beach Geomarket remained predominantly White (62%). Remanining Geomarkets exhibited a more racially/ethnic diverse population by 2020, including San Fernando Valley-West- (44% White, 4% Black, 39% Hispanic; 11% Asian Pacific Islander), San Fernando Valley -East (43% White, 4% Black, 41% Hispanic; 8% Asian Pacific Islander), Glendale and Pasadena (35% White, 3% Black, 37% Hispanic; 21% Asian Pacific Islander), South Bay (29% White, 12% Black, 33% Hispanic; 21% Asian Pacific Islander), Long Beach (26% White, 11% Black, 46% Hispanic; 13% Asian Pacific Islander), Covina and West Covina (19% White, 3% Black, 52% Hispanic; 24% Asian Pacific Islander), and Riverside, San Bernardino, and Ontario (28% White, 7% Black, 54% Hispanic; 6% Asian Pacific Islander). However, the South and South Central Los Angeles and Hollywood and Wilshire Geomarkets experienced substantial declines in the share of Black residents from 1980 to 2020, dropping from 44% to 18% and 18% to 9%, respectively.

[I THINK THE LEADS ARE THAT SOUTH CENTRAL (ESPECIALLY), EAST LA, AND WILSHIRE ARE THE POOREST. START TEHRE AND THEN GET TO OTHER RESULTS.]
@fig-la-rq1-ses presents trends in socioeconomic indicators for the Los Angeles metropolitan area from 1980 to 2020. While median income increased across all of the metropolitan area over the three census periods, some shifts occurred in the relative socioeconomic levels across Geomarkets. Many Geomarkets kept their relative levels of affluence from 1980 to 2020. For instance, the Covina and West Covina, South Bay, and San Fernando Valley remained among the most affluent, whereas the San Fernando Valley East, Glendale and Pasadena, Long Beach, and Riverside, San Bernardino, and Ontario Geomarkets remained at middle ranges of affluence and the Hollywood and Wilshire, South and South Central Los Angeles, and East Los Angeles Geomarkets remained at the lowest levels of affluence across the metropolitan area during this period. The West Los Angeles and West Beach Geomarket demonstrated the most substantial increase in median household income, rising from \$112,000 in 1980 to \$132,000 in 2000, and reaching \$137,000 by 2020. This sustained growth positioned it as the most affluent Geomarket in the Los Angeles metropolitan area.

@@

[THINK YOU CAN CUT THIS PARAGRAPH] The Los Angeles metropolitan area exemplifies a complex multi-nuclei urban model, characterized by decentralized population growth and increasing demographic diversity across its 11 Geomarkets. From 1980 to 2020, the region’s population expanded significantly leading to overall declines in the White population and increases in Hispanic and Asian populations across Geomarkets. West Los Angeles and West Beach is the only majority White Geomarket in 2020. Socioeconomic indicators reveal enduring disparities, especially in historically lower-income Geomarkets like South and South Central Los Angeles, East Los Angeles, and Hollywood and Wilshire. Meanwhile, Geomarkets such as West Los Angeles and West Beach became increasingly affluent and highly educated, highlighting growing socioeconomic and educational divides. Given these demographic shifts within the context of a multi-nuclei urban model, the Los Angeles metropolitan area presents a distinctive Geomarket landscape case in which racial/ethnic enclaves across a range of various socioeconomic levels are distributed across multiple Geomarkets rather than concentrated in one or two.

@@

To examine how the socioeconomic and racial composition of included versus excluded prospects varies when student list purchases filter on particular Geomarkets in the Los Angeles metropolitan areas, we analyze six orders placed by a research university that filtered for prospects across the entire state of California, in the 2019-2020 [NOT JUST 2019-20. REVISE] high school graduating class, and by PSAT scores. Two orders indicated PSAT score thresholds ranging from 1070-1180, another two orders filtered for a 1190-1260 PSAT range, and the remaining two orders filtered for scores ranging from 1270-1520.

>>>>

<!--# DISCUSSION OF THIS FIGURE SOMEWHAT BURIES THE LEAD THAT CHICAGO LOOKS TOTALLY DIFFERENT THAN THE OTHERS IN TERMS OF RACIAL COMPOSITION OF PURCHASED PROFILES! -->

@fig-chicago-rq2-race presents the racial/ethnic composition of purchased student profiles within Geomarkets in the Chicago metropolitan area, notably revealing that the City of Chicago stands out distinctly from all other Geomarkets in terms of the substantial racial and ethnic diversity of its purchased profiles. At the lowest PSAT score thresholds (top panel, 1020-1150) the racial/ethnic distribution of prospects within Geomarkets most closely reflects the demographic makeup of the metropolitan area in 2020 (see @fig-chicago-rq1-race). For example, more than 8,000 of the 15,981 total prospects in this SAT range order were concentrated in the Western Suburbs (56% White, 8% Asian, 6% Black, 26% Hispanic, 4% Multiracial) and South and Southwest Suburbs (55% White, 3% Asian, 17% Black, 22% Hispanic, and 3% Multiracial), both of which yield majority White prospect pools comparable to the racial/ethnic makeup of their overall Geomarket populations. The North Shore (78% White, 8% Asian, 2% Black, 8% Hispanic) and the Northwest Suburbs (61% White, 11% Asian, 2% Black, 23% Hispanic) Geomarket pools exhibit the greatest proportions of White prospects. Unlike any other Geomarket, nearly nine of every 10 prospects from the City of Chicago Geomarket were Hispanic (50%), Black (30%), Asian (5%), or Multiracial (2%).

@@

<!--#  [WRITING IS QUITE GOOD!I THINK THERE IS A MODEST LEAD THAT THE NEXT COUPLE PARAGRAPHS DON'T STRESS ENOUGH , WHICH IS THAT TX19/CITY OF DALLAS IS THE MOST LEAST WHITE/ASIAN AND THE POOREST]--> @fig-dallas-rq1-race presents the racial and ethnic composition of each Geomarket within the Dallas-Fort Worth metropolitan area across the three census periods. In 1980, all Geomarkets were predominantly White, with White populations ranging from 62% in City of Dallas to nearly 92% in West of Dallas/Fort Worth Metroplex. By 2000, however, Geomarkets located in the central part of the metropolitan area exhibited a more balanced racial/ethnic distribution, including Dallas County (58% White, 16% Black, 19% Hispanic; 6% Asian), Irving, Arlington, and Grand Prairie (54% White, 12% Black, 26% Hispanic; 16% Asian), and City of Fort Worth (57% White, 16% Black, 22% Hispanic; 3% Asian). In contrast, Collin & Rockwall Counties and West of Dallas/Fort Worth Metroplex Geomarkets remained predominantly White in 2000 at 78% and 80%, respectively.

By 2000, the City of Dallas was no longer a predominantly White Geomarket, with a population that was approximately 24% Black, 30% Hispanic, and 3% Asian. Over the next two decades, the Hispanic population grew steadily, contributing to greater diversity. By 2020, the City of Dallas Geomarket had the lowest proportions of White and Asian residents across the Dallas-Fort Worth Metropolitan area, with approximately 21% Black, 38% Hispanic, and 2% Asian residents.

These shifts in racial and ethnic composition also became more pronounced for other Geomarkets between 2000 and 2020. The Collin and Rockwall and West of Dallas/Fort Worth Metroplex Geomarkets experienced continued declines in their share of White residents. However, both Geomarkets remained predominantly White (58% and 63%, respectively). In contrast, all other Geomarkets experienced continued increases in the shares of Black and Hispanic residents, resulting in a diminished previously White-majority populations for Dallas County (33% White, 23% Black, 32% Hispanic; 9% Asian), Irving, Arlington, and Grand Prairie (30% White, 20% Black, 36% Hispanic; 12% Asian), and City of Fort Worth (46% White, 15% Black, 33% Hispanic; 4% Asian). <!--#  Lastly, continuous increases in the shares of Black and Hispanic residents further diversified the City of Dallas, making it the most racially/ethnically diverse Geomarket in the Dallas-Fort Worth metropolitan area. -->

<!--# [CITY OF DALLAS!] --> Socioeconomic demographics of median income, percent living below poverty, and educational attainment over time are presented in @fig-dallas-rq1-ses for the Dallas-Fort Worth metropolitan area. Overall, median income both within and across Geomarkets remained relatively stable across the three census periods. Between 1980 to 2000, affluence levels across Geomarkets were consistent, ranging from a low of \$82,000 in City of Fort Worth to a high of \$135,00 in the Collin & Rockwall Counties Geomarket. However, the substantial increase in median income for the Collin & Rockwall Counties Geomarket during this time period plateaued by 2020, reshuffling relatively affluence levels among Geomarkets. By 2020, Collin & Rockwall Counties remained the most affluent Geomarket with a median household income of \$133,000, followed by West of Dallas/Fort Worth Metroplex (\$116,000), and Dallas County (\$90,000).

In contrast, median household income for City of Dallas and City of Fort Worth remained relatively stable through 2020. As a result, both City of Dallas and City of Fort Worth continued to rank at the lowest levels of affluence within the metropolitan area at \$86,000 and \$87,000 median household incomes, respectively. @fig-dallas-rq1-ses also illustrates changes in the percentage of residents living in poverty and educational attainment within Geomarkets over time, revealing similar dynamics across Geomarkets to those observed in median household income. Together, these trends underscore the growingly racial/ethnic diverse but widening economic inequality within the evolving Geomarket landscape of the Dallas-Fort Worth metroplex.

<!--#  [DON'T KNOW IF WE NEED PARAGRAPH ON POVERTY + EDUCATIONAL ATTAINMENT] @fig-dallas-rq1-ses also illustrates changes in the percentage of residents living in poverty within Geomarkets over time, revealing shifting dynamics across Geomarkets similar to those observed in median household income. Together, these trends underscore the growingly racial/ethnic diverse but widening economic inequality within the evolving Geomarket landscape of the Dallas-Fort Worth metroplex. In 1980, poverty rates were comparably low in both Collin & Rockwall Counties and Dallas County, moderate in the Irving, Arlington, and Grand Prarie and West of Dallas/Fort Worth Metroplex Geomarkets, and highest for City of Dallas and City of Fort Worth Geomarkets. Overtime, poverty rates declined in Collin & Rockwall Counties and West of Dallas/Fort Worth Metroplex, while increasing for all other Geomarkets. These divergent trends resulted in a widening gap in poverty levels across the Dallas-Fort Worth metropolitan area. A similar widening disparity is evident in educational attainment. Although all Geomarkets experienced increases in the percentage of residents with a bachelor's degree or higher between 1980 and 2020, the Collin & Rockwall Counties Geomarket experienced a substantially greater increase - from 23% in 1980 to 50% in 2020- maintaining the gap with other Geomarkets in terms of educational attainment. .-->

<!--#  [DON'T THINK YOU NEED SUMMARY PARAGRAPHS LIKE THIS WITHIN SECTIONS FOR THIS METRO]The Dallas-Fort Worth metroplex experienced substantial demographic growth since Geomarket boundaries were dilineated by @RN4982 in the early 1980s, leading to an urban landscape with considerable racial/ethnic diversity across multiple Geomarkets (as opposed to concentrated within a single Geomarket like Chicago) but stark socioeconomic disparities between Geomarkets. While all six Geomarkets experienced population increases, the most significant growth occurred in the outer suburban areas, reflecting ongoing urban sprawl. Such growth is largely attributed to increases in Hispanic, Black, and Asian populations leading to only two of the six Dallas-Fort Worth Geomarkets remaining predominantly White (Collin and Rockwall Counties and West of Dallas/Fort Worth Metroplex. On the other hand, the metropolitan area has experienced widening disparities in income, poverty, and education attainment between Geomarkets, especially between the historically affluent Collin and Rockwall Counties and the less economically advantaged inner-core Geomarkets (City of Fort Worth and City of Dallas). Together, these trends underscore the growingly racial/ethnic diverse but widening economic inequality within the evolving Geomarket landscape of the Dallas-Fort Worth metroplex. -->

@@

Contributions of each Geomarket to the pool of prospects based on first-generation college student status from the Dallas-Fort Worth metropolitan area are presented in @fig-dallas-rq2-midSAT-ses. Only 303 prospects of the 3,928 total in this middle PSAT score range pool were first-generation college students whose parents did not attend college. Again, all Geomarkets besides Collin and Rockwall Counties (14% versus 36% for all prospects) and West of Dallas/Fort Worth Metroplex (15% versus 24% for all prospects) contributed equal or larger proportions of first-generation college students (no college) relative to their contributions to the overall prospect pool. Disproportioantely larger contributions range from six percentage points by the Dallas Counties Geomarket (19% versus 13% for all prospects) to 11 percentage points by the City of Dallas Geomarket (22% versus 11% for all prospects). Contributions become more proportional for the pool of first-generation college students whose parents attended but did not complete their degree. Proportional balances across Geomarkets are also evident for the 3,069 prospects that are not first-generation college students, with the exception of Collin and Rockwall Counties Geomarket contributing the largest disproportionate share of not first-generation college students (40%) relative to the contributions to the overall prospect pool (36%).

@@

By contrast, for non-White groups, a larger share of prospects were first-generation students. Furthermore, first generation status differed substantially across Geomarkets, especially for Hispanic and Asian students. For Hispanic students, more than 50% of prospects from the City of Dallas, City of Forth Worth, Arlington & Grand Prairie, and Dallas County are first-generation college students. By contrast, for the more affluent Collin & Rockwall Geomarket and the West of Dallas/Fort Worth Metroplex Geomarkets, less than 30% of prospects were first-generation. For Asian prospects, City of Fort Worth and Irving, Arlington, & Grand Prairie Geomarkets contributed disproportionately larger shares of first-generation college students across both parents with no college and some college. By contrast, the more affluent Collin & Rockwall sent more Asian prospects (285) than any other Geomarket, but about 87% of these prospects were not first-generation college students.

```{=html}
<!-- I ADDED THE ABOVE PARAGRAPH AND CUT THIS PARAGRAPH TO GIVE THE LEDE MORE PUNCH 
Similar patterns of one or two Geomarkets disproportionately contributing larger shares of first-generation college students was evident across other racial/ethnic categories in @fig-dallas-rq2-midSAT-combo. For instance, City of Dallas (18%) and City of Fort Worth (24%) contributed larger proportions of first-generation college students whose parents did not attend college relative to the overall pool of all prospects (8%). On the other hand, the Irving, Arlington, & Grand Prairie (23%) and Dallas Counties (26%) Geomarkets contributed disproportionately larger proportions of Black students whose parents attended some college but did not complete their degree relative to the overall pool of all prospects (17%). For Asian prospects, City of Fort Worth and Irving, Arlington, & Grand Prairie Geomarkets contributed disproportionately larger shares of first-generation college students across both parents with no college and some college. City of Dallas (38%), City of Fort Worth (39%), Irving, Arlington, & Grand Prairie (32%), Dallas Counties (28%) all contributed larger shares of Hispanic students whose parents did not attend college relative to the overall prospect pool of these first-generation college students (23%), whereas only Dallas Counties contributed to larger shares of Hispanic students who parents attended some college but did not attain their degree (26% versus 20% for all prospects).
-->
```
Analyses of student list orders across the Dallas-Fort Worth metropolitan area reveals patterns of racial/ethnic and socioeconomic stratification in Geomarket contributions. As Dallas-Fort Worth became increasingly diverse from 1980 to 2020, a more balanced racial/ethnic distribution emerged across the metropolitan area's Geomarket landscape. However, these patterns did not translate to racial/ethnic composition of student prospect pools. Suburban Geomarkets like Collin and Rockwall Counties and the West of Dallas/Fort Worth Metroplex yielded a disproportionately larger share of Asian prospects, whereas central Geomarkets such as the City of Dallas, Dallas County, and Irving, Arlington, and Grand Prairie contributed a disproportionately larger share of Hispanic and Black prospects. These racial/ethnic disparities become significantly more pronounced, particularly in the disproportionate share of Asian prospects, as PSAT score thresholds increase (see Appendix @fig-dallas-rq2-lowSAT-race and @fig-dallas-rq2-highSAT-race). Similar Geomarket contribution patterns are also evident by first-generation college student status, with Collin and Rockwall Counties and the West of Dallas/Fort Worth Metroplex providing larger shares of non-first generation college student prospects, and central Geomarkets (City of Dallas, City of Fort Worth, Dallas County, and Irving, Arlington, and Grand Prairie) providing disproportionately larger shares of first-generation college student prospects. Moreover, central Geomarkets are particularly important for including Black (City of Dallas and City of Fort Worth), Hispanic (City of Dallas, City of Fort Worth, Dallas County, and Irving, Arlington, and Grand Prairie), and Asian (City of Fort Worth and Irving, Arlington, and Grand Prairie) first-generation college students. While Collin and Rockwall Counties and the West of Dallas/Fort Worth Metroplex Geomarkets still contribute a notable overall number of Black, Hispanic, and Asian prospects, such students are predominantly not first-generation college students.

@@

\@fig-la-rq1-race presents the racial and ethnic composition of each Geomarket within the Los Angeles metropolitan area from 1980 to 2020. Over this period, the proportion of White residents declined across all Geomarkets, while Asian and Hispanic populations generally increased. \[THE CAVEAT THAT WE DON'T HAVE ASIAN FOR 1980 IS MORE IMPORTANT FOR LA\] In 1980, all Geomarkets - except for East Los Angeles, South and South Central Los Angeles, and Hollywood and Wilshire- were predominantly White, with percentages ranging from 62% White in Long Beach to 78% in the San Fernando Valleys (West and East) Geomarkets. By 2020, however, only the West Los Angeles and West Beach Geomarket remained predominantly White (62%). Remanining Geomarkets exhibited a more racially/ethnic diverse population by 2020, including San Fernando Valley-West- (44% White, 4% Black, 39% Hispanic; 11% Asian Pacific Islander), San Fernando Valley -East (43% White, 4% Black, 41% Hispanic; 8% Asian Pacific Islander), Glendale and Pasadena (35% White, 3% Black, 37% Hispanic; 21% Asian Pacific Islander), South Bay (29% White, 12% Black, 33% Hispanic; 21% Asian Pacific Islander), Long Beach (26% White, 11% Black, 46% Hispanic; 13% Asian Pacific Islander), Covina and West Covina (19% White, 3% Black, 52% Hispanic; 24% Asian Pacific Islander), and Riverside, San Bernardino, and Ontario (28% White, 7% Black, 54% Hispanic; 6% Asian Pacific Islander). However, the South and South Central Los Angeles and Hollywood and Wilshire Geomarkets experienced substantial declines in the share of Black residents from 1980 to 2020, dropping from 44% to 18% and 18% to 9%, respectively.

\[I THINK THE LEADS ARE THAT SOUTH CENTRAL (ESPECIALLY), EAST LA, AND WILSHIRE ARE THE POOREST. START TEHRE AND THEN GET TO OTHER RESULTS.\] @fig-la-rq1-ses presents trends in socioeconomic indicators for the Los Angeles metropolitan area from 1980 to 2020. While median income increased across all of the metropolitan area over the three census periods, some shifts occurred in the relative socioeconomic levels across Geomarkets. Many Geomarkets kept their relative levels of affluence from 1980 to 2020. For instance, the Covina and West Covina, South Bay, and San Fernando Valley remained among the most affluent, whereas the San Fernando Valley East, Glendale and Pasadena, Long Beach, and Riverside, San Bernardino, and Ontario Geomarkets remained at middle ranges of affluence and the Hollywood and Wilshire, South and South Central Los Angeles, and East Los Angeles Geomarkets remained at the lowest levels of affluence across the metropolitan area during this period. The West Los Angeles and West Beach Geomarket demonstrated the most substantial increase in median household income, rising from \$112,000 in 1980 to \$132,000 in 2000, and reaching \$137,000 by 2020. This sustained growth positioned it as the most affluent Geomarket in the Los Angeles metropolitan area.

@@

\[THINK YOU CAN CUT THIS PARAGRAPH\] The Los Angeles metropolitan area exemplifies a complex multi-nuclei urban model, characterized by decentralized population growth and increasing demographic diversity across its 11 Geomarkets. From 1980 to 2020, the region’s population expanded significantly leading to overall declines in the White population and increases in Hispanic and Asian populations across Geomarkets. West Los Angeles and West Beach is the only majority White Geomarket in 2020. Socioeconomic indicators reveal enduring disparities, especially in historically lower-income Geomarkets like South and South Central Los Angeles, East Los Angeles, and Hollywood and Wilshire. Meanwhile, Geomarkets such as West Los Angeles and West Beach became increasingly affluent and highly educated, highlighting growing socioeconomic and educational divides. Given these demographic shifts within the context of a multi-nuclei urban model, the Los Angeles metropolitan area presents a distinctive Geomarket landscape case in which racial/ethnic enclaves across a range of various socioeconomic levels are distributed across multiple Geomarkets rather than concentrated in one or two.

@@

To examine how the socioeconomic and racial composition of included versus excluded prospects varies when student list purchases filter on particular Geomarkets in the Los Angeles metropolitan areas, we analyze six orders placed by a research university that filtered for prospects across the entire state of California, in the 2019-2020 \[NOT JUST 2019-20. REVISE\] high school graduating class, and by PSAT scores. Two orders indicated PSAT score thresholds ranging from 1070-1180, another two orders filtered for a 1190-1260 PSAT range, and the remaining two orders filtered for scores ranging from 1270-1520.

@fig-la-rq2-race presents the racial/ethnic composition of purchased student profiles. The more than 15,000 prospects in the resulting student lists at the low PSAT score thresholds (1070-1180) reflect very similar racial/ethnic compositions within Geomarkets than overall demographic makeup of the metropolitan area in 2020 (see @fig-la-rq1-race). \[THE LEDE IS SOUTH CENTRAL AND EAST LA. START THERE!\]. However, the racial/ethnic composition of prospects become less proportional to overall metropolitan patterns by Geomarket as PSAT scores increase. For instance, prospect pools at middle range PSAT scores (1190-1260) within Geomarkets become disproportionately more Asian, while the representation of Hispanic and Black prospects decline substantially. This pattern becomes most pronounced at the highest PSAT score prospect pools. \[I DON'T THINK THAT BLACK AND LATINX STUDENTS HAVING LOW TEST SCORES IS THE STORY WE ARE TRYING TO TELL. THAT STORY HAS BEEN TOLD. THE MORE IMPORTANT THING IS, WITHIN SOME TEST THRESHOLD, HOW DOES THE RACIAL COMPOSITION VARY ACROSS GEOMARKETS. LIKE EVEN WITHIN THE MID RANGE, SOUTH CENTRAL, EAST LA, AND INLAND EMPIRE STAND OUT\] <!-- This disproportionate over representation for Asian prospects in comparison to the population of residents with Geomarkets ranges from 3 percentage points in the West Los Angeles and West Beach Geomarket (16% prospects versus 13% Asian residents) to 52 percentage points in the East Los Angeles Geomarket (79% prospects versus 27% residents). @fig-la-rq2-race illustrates the share of Black prospects within these high PSAT orders declines to less than 3% within all Geomarkets in comparison to Black residents making up a range of 3% (Covina and West Covina) to 18% (South and South Central Los Angeles) of residents within Geomarkets. Similar underrepresentation patterns are evident for Hispanic porspects are evident across all Geomarkets execept for a nearly equal representation in South and South Central Los Angeles (73% prospects versus 71% Hispanic residents). -->
